---
title: 'Figure 3: PCA; Changes in Up- and Downregulated syllables'
format:
    html:
        code-fold: false
        self-contained: true
        toc: true
editor:
    render-on-save: false
---

Figure 3 demonstrates that female mice show behavioral variability distinct of that observed in males as a result of CSW/DS, and that this behavioral variability is driven by an attenuated behavioral regulation in response to target presence. This code generates Fig 3 and Fig. S2 as well as statistics for the results section **Dimensionality reduction identifies distinct sources of stress-related variance in males and females**.

### Libraries & Config

```{r}
#| label: libraries
#| message: false
#| code-fold: true

library(here)
library(glue)
library(dplyr)
library(tidyr)
library(knitr)
library(pander)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(cowplot)
```

```{r}
#| label: config + paths + data
#| code-fold: true

source(here('styles.R'))
data_path <- here('data')
SIT_diff <- readRDS(glue('{data_path}/SIT_diff_primary.rds'))
SIT_z_diff <- readRDS(glue('{data_path}/SIT_z_diff_primary.rds'))
syl_info <- tibble(read.csv(glue('{data_path}/syllable_info.csv')))

```

### Run PCA

We first ran a PCA on SI ratio, corner ratio and the NT - T difference for each syllable. We looked exclusively at the first two principle components.

```{r}
#| label: pca
#| code-fold: true
#| warning: false

# Include IZ Ratio, Corner Ratio, and all syllables
SIT_pca <- SIT_diff[,c(grep('_time', names(SIT_diff)),
                       grep('syl', names(SIT_diff)))]
pca <- prcomp(SIT_pca, scale = TRUE, center = TRUE) #run PCA

# Add PCA results to full data frame
SIT_diff$PC1 <- pca$x[,1]
SIT_diff$PC2 <- pca$x[,2]

# Get factor loadings and prep for plotting
SIT_pca <- tibble(var = names(SIT_pca), 
                  PC1 = as.numeric(pca$rotation[,1]), 
                  PC2 = as.numeric(pca$rotation[,2]))
SIT_pca$new_label <- syl_info$new_label[match(SIT_pca$var, syl_info$metric)]
SIT_pca$avg_freq <- syl_info$avg_freq[match(SIT_pca$var, syl_info$metric)]
SIT_pca$new_label <- ifelse(is.na(SIT_pca$new_label), 
                            gsub('_time', ' Ratio', SIT_pca$var), 
                            SIT_pca$new_label)
SIT_pca$new_label[SIT_pca$new_label == 'IZ Ratio'] = 'SI Ratio'
SIT_pca$social_identity <- ifelse(SIT_pca$new_label %in% c('SI Ratio', 'Corner Ratio'), 
                                  'A', substr(SIT_pca$new_label, 1, 1))

# Calcualte PC1 and PC2 mean for plotting
SIT_diff %>%
  group_by(Sex, Condition, full_cond) %>%
  summarise(avg_PC1 = mean(PC1), avg_PC2 = mean(PC2),
            sd_PC1 = sd(PC1), sd_PC2 = sd(PC2)) -> SIT_pca_mean
```

------------------------------------------------------------------------

### PCA Plots (Fig. 3a, 3b)

The PC plot shows that CSW/DS induces behavioral variability largely captured along PC1 in males, but along PC2 in females. This analysis not only confirms that male and female mice differ in their behavioral response after CSW/DS in the SIT, but also emphasizes that CSW/DS indeed causes behavioral changes in witness females, disputing the hypothesis that the stress model does not work.

```{r}
#| label: Fig 3a prep and plot
#| fig-width: 10
#| fig-asp: 0.5
#| fig-align: center
#| code-fold: true
#| warning: false
#| error: false

# Female plots
F_pca <- ggplot(SIT_diff[SIT_diff$Sex == 'Female',]) + 
           geom_hline(yintercept = 0) + 
           geom_vline(xintercept = 0) + 
           geom_point(aes(PC1, PC2, color=full_cond), 
                          size = 2, alpha = 0.3) + 
           geom_point(data = SIT_pca_mean[SIT_pca_mean$Sex == 'Female',], 
                      aes(avg_PC1, avg_PC2, 
                          color=full_cond), size = 5) +
           ylim(c(-4,4)) + xlim(c(-4,4)) + 
           scale_color_manual(values = col_full[c(1,2)]) + 
           plot_theme + 
           theme(legend.position = 'none',
                 plot.margin = unit(c(0,0,0,0), 'cm'))

F_hist_pc1 <- ggplot(SIT_diff[SIT_diff$Sex == 'Female',], 
                     aes(PC1, color=full_cond, fill = full_cond)) + 
                geom_hline(yintercept = 0) +
                geom_density(alpha = 0.5) + 
                xlim(c(-4,4)) + 
                scale_color_manual(values = col_full[c(1,2)]) + 
                scale_fill_manual(values = col_full[c(1,2)]) + 
                plot_theme + 
                theme(legend.position = 'none',
                      plot.margin = unit(c(0,0,0,0), 'cm')) + 
                clean_theme()

F_hist_pc2 <- ggplot(SIT_diff[SIT_diff$Sex == 'Female',], 
                     aes(PC2, color=full_cond, fill = full_cond)) + 
                geom_hline(yintercept = 0) +
                geom_density(alpha = 0.5) + 
                xlim(c(-4,4)) + 
                scale_color_manual(values = col_full[c(1,2)]) + 
                scale_fill_manual(values = col_full[c(1,2)]) + 
                coord_flip() +
                plot_theme + 
                theme(legend.position = 'none',
                      plot.margin = unit(c(0,0,0,0), 'cm')) + 
                clean_theme()

F_plot <- plot_grid(F_hist_pc1, NULL, F_pca, F_hist_pc2, align = 'hv', 
                    rel_widths = c(3,1), rel_heights = c(1,3))

# Male plots
M_pca <- ggplot(SIT_diff[SIT_diff$Sex == 'Male',]) + 
           geom_hline(yintercept = 0) + 
           geom_vline(xintercept = 0) + 
           geom_point(aes(PC1, PC2, color=full_cond), 
                          size = 2, alpha = 0.3) + 
           geom_point(data = SIT_pca_mean[SIT_pca_mean$Sex == 'Male',], 
                      aes(avg_PC1, avg_PC2, 
                          color=full_cond), size = 5) +
           ylim(c(-4,4)) + xlim(c(-4,4)) + 
           scale_color_manual(values = col_full[c(3,4)]) + 
           plot_theme + 
           theme(legend.position = 'none',
                 plot.margin = unit(c(0,0,0,0), 'cm'))

M_hist_pc1 <- ggplot(SIT_diff[SIT_diff$Sex == 'Male',], 
                     aes(PC1, color=full_cond, fill = full_cond)) + 
                geom_hline(yintercept = 0) +
                geom_density(alpha = 0.5) + 
                xlim(c(-4,4)) + 
                scale_color_manual(values = col_full[c(3,4)]) + 
                scale_fill_manual(values = col_full[c(3,4)]) + 
                plot_theme + 
                theme(legend.position = 'none',
                      plot.margin = unit(c(0,0,0,0), 'cm')) + 
                clean_theme()

M_hist_pc2 <- ggplot(SIT_diff[SIT_diff$Sex == 'Male',], 
                     aes(PC2, color=full_cond, fill = full_cond)) + 
                geom_hline(yintercept = 0) +
                geom_density(alpha = 0.5) + 
                xlim(c(-4,4)) + 
                scale_color_manual(values = col_full[c(3,4)]) + 
                scale_fill_manual(values = col_full[c(3,4)]) + 
                coord_flip() +
                plot_theme + 
                theme(legend.position = 'none',
                      plot.margin = unit(c(0,0,0,0), 'cm')) + 
                clean_theme()

M_plot <- plot_grid(M_hist_pc1, NULL, M_pca, M_hist_pc2, align = 'hv', 
                    rel_widths = c(3,1), rel_heights = c(1,3))

# Plot 
plot_grid(M_plot, F_plot)
```

The factor loadings of the PC plot show that PC1 is mostly driven by location-based metrics, i.e., SI ratio and corner ratio. In contrast, PC2 is driven by syllable frequency differences. Notably, Up- and Downregulated syllables have opposing effects on PC2.

```{r}
#| label: Fig 3b prep and plot
#| code-fold: true
#| fig-width: 5
#| fig-asp: 1
#| fig-align: center
#| warning: false
#| error: false

ggplot(SIT_pca, aes(PC1, PC2, color = social_identity, fill = social_identity)) + 
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) + 
  geom_label_repel(aes(label = new_label), 
                   size = 5, alpha = 0.2, seed = 1234 , fontface = 'bold')+ 
  geom_label_repel(aes(label = new_label), 
                   size = 5, alpha = 1, seed = 1234,  
                   label.size = NA, fill = NA, fontface = 'bold') + 
  scale_color_manual(values = col_syls) + 
  scale_fill_manual(values = col_syls) + 
  ylim(c(-0.6,0.6)) + xlim(c(-0.8,0.8)) + 
  plot_theme + 
  clean_theme() + theme(legend.position = 'none')
```

### Importance of Syllable Social Identity (Fig. 3c, 3d)

Given that Up- and Downregulated syllables split along the axis of PC2, which in turn differentiates control and stressed females, we next looked at the average changes in syllable frequency observed within each social identity group. We show that in males, Upregulated syllables are more upregulated in defeated males compared to controls. In females, we see an attenuated regulation in both Up- and Downregulated syllables.

We summarize these observations by quantifying the average absolute syllable frequency change. We observed an increase in syllable frequency

```{r}
#| label: Fig 3cd prep and plot
#| code-fold: true
#| fig-width: 10
#| fig-asp: 2
#| fig-align: center
#| warning: false
#| error: false

# Annotate syllables with social identity
N_syls <- syl_info$metric[grepl('N',syl_info$new_label)]
U_syls <- syl_info$metric[grepl('U',syl_info$new_label)] 
D_syls <- syl_info$metric[grepl('D',syl_info$new_label)]

SIT_z_diff$Social <- factor(ifelse(SIT_z_diff$metric %in% N_syls, 'No Change', 
                            ifelse(SIT_z_diff$metric %in% U_syls, 'Up', 'Down')),
                            levels = c('Up', 'No Change', 'Down'))

# Calculate mean per social identity type and abs change
SIT_z_diff %>%
  group_by(id_full, full_cond, Sex, Condition, Social, cohort) %>%
  summarise(Social_avg = mean(z_diff)) -> SIT_syl_avg_freq_z_diff

SIT_z_diff %>%
  group_by(id_full, full_cond, Sex, Condition, Social, cohort) %>%
  mutate(all = 'All Syllables') %>% # So that plot can be titled by facet
  summarise(Avg_diff = mean(abs(z_diff))) -> SIT_avg_freq_z_diff

# Plot for each social identity type
### Not faceted because of different y axes
U_plot <- ggplot(SIT_syl_avg_freq_z_diff[SIT_syl_avg_freq_z_diff$Social == 'Up',], 
                 aes(Condition, Social_avg, color = full_cond, fill = full_cond)) + 
            geom_hline(yintercept = 0) + 
            geom_jitter(width = 0.2, alpha = 0.2, height = 0) +
            stat_summary(fun.y = 'mean', geom = 'point', size = 2) + 
            stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                         linewidth = 1, width = 0.4) + 
            scale_y_continuous(name = 'Mean Syllable Frequency Diff. (Z Score)', 
                               limits = c(-10,10), breaks = c(-10,0,10)) + 
            facet_grid(Sex ~ Social, scale = 'free_y') + 
            scale_color_manual(values = col_full[c(1,3,2,4)]) + 
            scale_fill_manual(values = col_full) + 
            plot_theme + 
            theme(legend.position = 'None',
                  axis.title.x = element_blank())

N_plot <- ggplot(SIT_syl_avg_freq_z_diff[SIT_syl_avg_freq_z_diff$Social == 'No Change',], 
                 aes(Condition, Social_avg, color = full_cond, fill = full_cond)) + 
            geom_hline(yintercept = 0) + 
            geom_jitter(width = 0.2, alpha = 0.2, height = 0) +
            stat_summary(fun.y = 'mean', geom = 'point', size = 2) + 
            stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                         linewidth = 1, width = 0.4) + 
            scale_y_continuous(name = 'Mean Syllable Frequency Diff. (Z Score)', 
                     limits = c(-5,5), breaks = c(-5,0,5)) + 
            facet_grid(Sex ~ Social, scale = 'free_y') + 
            scale_color_manual(values = col_full[c(1,3,2,4)]) + 
            scale_fill_manual(values = col_full) + 
            plot_theme + 
            theme(legend.position = 'None',
                  axis.title.x = element_blank())

D_plot <- ggplot(SIT_syl_avg_freq_z_diff[SIT_syl_avg_freq_z_diff$Social == 'Down',], 
                 aes(Condition, Social_avg, color = full_cond, fill = full_cond)) + 
            geom_hline(yintercept = 0) + 
            geom_jitter(width = 0.2, alpha = 0.2, height = 0) +
            stat_summary(fun.y = 'mean', geom = 'point', size = 2) + 
            stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                         linewidth = 1, width = 0.4) + 
            scale_y_continuous(name = 'Mean Syllable Frequency Diff. (Z Score)', 
                     limits = c(-2,2), breaks = c(-2,0,2)) + 
            facet_grid(Sex ~ Social, scale = 'free_y') + 
            scale_color_manual(values = col_full[c(1,3,2,4)]) + 
            scale_fill_manual(values = col_full) + 
            plot_theme + 
            theme(legend.position = 'None',
                  axis.title.x = element_blank())

SMS_plot <- ggplot(SIT_avg_freq_z_diff, 
                   aes(Condition, Avg_diff, color = full_cond, fill = full_cond)) + 
              geom_jitter(width = 0.2, alpha = 0.2, height = 0) +
              stat_summary(fun.y = 'mean', geom = 'point', size = 2) + 
              stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                           linewidth = 1, width = 0.4) + 
              ylab('Social Modulation Score (SMS)') + 
              facet_grid(Sex ~ all) + 
              scale_color_manual(values = col_full) + 
              scale_fill_manual(values = col_full) + 
              plot_theme + 
              theme(legend.position = 'None',
                    axis.title.x = element_blank())

plot_grid(U_plot, N_plot, D_plot, NULL, SMS_plot, align = 'hv', nrow = 1)
```

```{r}
#| label: Fig 3cd stats
#| code-fold: true
#| warning: false
#| error: false

```

### SI Ratio and SMS correlation to PCs (Fig. 3e)

Finally, we confirm that SI ratio correlates with PC1 and that SMS indeed captures the variability identified by PC2.

```{r}
#| label: Fig 3e plot
#| code-fold: true
#| fig-width: 10
#| fig-asp: 0.5
#| fig-align: center
#| warning: false
#| error: false

pc1 <- ggplot(SIT_diff, aes(PC1, IZ_time)) + 
         geom_point(aes(color=full_cond), size = 2, alpha = 0.3) + 
         stat_poly_line(col = 'black') +
         stat_poly_eq(size = 7, label.y = 'top', label.x = 'right') +
         scale_color_manual(values = col_full) + 
         xlab('PC1') + ylab('SI Ratio') +
         xlim(c(-4,4)) + 
         plot_theme + 
         theme(legend.position = 'none')

pc2 <- ggplot(SIT_diff, aes(PC2, Diff_score)) + 
         geom_point(aes(color=full_cond), size = 2, alpha = 0.3) + 
         stat_poly_line(col = 'black') +
         stat_poly_eq(size = 7, label.y = 'top', label.x = 'right') +
         scale_color_manual(values = col_full) + 
         xlab('PC2') + ylab('SMS') +
         xlim(c(-4,4)) + 
         plot_theme + 
         theme(legend.position = 'none')

plot_grid(pc1, pc2, ncol = 2, nrow = 1, align = 'hv')
```

------------------------------------------------------------------------

### Defeat male-specific behavior: U1 Upregulation (Fig. S2)

We observed that Males show a higher upregulation of upregulated syllables only. When further investigating this observation we see that this is driven by an increase specific to U1 in stressed Males compared to control males.

```{r}

```

We then explored whether there mice express this behavior preferentially at specific locations in the arena and whether it is correlated with SI ratio. Indeed, mice in an example cohort (E18) display this syllable specifically around the interaction zone and in the corners, suggesting that it is a target induced behavior. Finally, we see that in stressed males, the frequency of U1 is correlated to SI ratio, such that animals with a lower SI ratio display a higher frequency of syllable U1.

```{r}

```

------------------------------------------------------------------------

## Session Info

```{r}
#| label: Session Info
#| echo: False
sessionInfo()
```