---
title: 'Figure 4: Velocity as a heuristic'
format:
    html:
        code-fold: false
        self-contained: true
        toc: true
editor:
    render-on-save: false
---

Figure 4 links the SMS to Velocity as a heuristic to quantify the female stress effect. This notebook generates Figure 4 and Figure S3 as well as statistics for the results section **Velocity is a unifying feature and heuristic for social modulation**.

```{r}
#| label: libraries
#| message: false
#| code-fold: true

library(here)
library(glue)
library(dplyr)
library(tidyr)
library(knitr)
library(scales)
library(nlme)
library(emmeans)
library(ggplot2)
library(cowplot)
```

```{r}
#| label: config + paths + data
#| code-fold: true

source(here('styles.R'))
data_path <- here('data')

SIT_diff <- readRDS(glue('{data_path}/SIT_diff_primary.rds'))
syl_char <- readRDS(glue('{data_path}/SIT_syl_descriptives.rds'))
syl_info <- tibble(read.csv(glue('{data_path}/syllable_info.csv')))
```

------------------------------------------------------------------------

### U and D syllables differ in motion and body pose characteristics (Fig. 4a)

```{r}
#| label: syl characteristics prep
#| code-fold: true
#| warning: false

# Annotate syl characteristics df with social identity
syl_char$syl_name = factor(syl_info$new_label[match(syl_char$syllable, 
                                                    substr(syl_info$metric, 4, 6))],
                           levels = c('U1', 'U2', 'N1', 'N2', 'D1', 'D2',
                                      'D3', 'D4', 'D5', 'D6', 'D7', 'D8', 'D9'))
syl_char$social_type = ifelse(grepl('U', syl_char$syl_name), 'Up',
                              ifelse(grepl('D', syl_char$syl_name), 'Down', 'No Change'))

# Avg per social identity type per animal
syl_char %<>%
  filter(social_type != 'No Change') %>% # Only interested in U and D
  mutate(outcome = ifelse(name %in% c('Velocity', 'Distance'), 'Motion', 
                          ifelse(name %in% c('Dist. to Center', 'Dist. to Target'),
                                             'Location', 'Pose'))) %>% # Group outcome for plotting
  group_by(cohort, id, id_full, name, syl_name, social_type, outcome) %>%
  summarise(mean = mean(mean)) %>% # Avg freq per syllable per animal 
  group_by(cohort, id, id_full, name, social_type, outcome) %>%
  summarise(mean = mean(mean)) %>% # Avg freq per social identity per animal 
  group_by(name) %>% #ungroup; group by characteristics
  mutate(z_mean = scale(mean)) %>% #scale and center for each characteristic
  na.omit()

# Stats: LME model & EM means contrast
lme_syl_char <- lme(z_mean ~ social_type * name, 
                    data = syl_char, 
                    random = ~ 1 | cohort/id)
lme_syl_char_table <- cbind(rownames(summary(lme_syl_char)$tTable),
                              round(unname(summary(lme_syl_char)$tTable[,'Value']), digits = 5),
                              round(unname(summary(lme_syl_char)$tTable[,'Std.Error']), digits = 5),
                              round(unname(summary(lme_syl_char)$tTable[,'DF']), digits = 5),
                              round(unname(summary(lme_syl_char)$tTable[,'t-value']), digits = 5))

mm_social <- emmeans(lme_syl_char, ~ social_type | name)
sidak_social_confint <- confint(pairs(mm_social, adjust = 'sidak')) # confint and emmeans for plots
tests <- tibble(name = sidak_social_confint$name,
                estimate = round(sidak_social_confint$estimate, digits = 5), 
                SE = round(sidak_social_confint$SE, digits = 5),
                lower_cl = round(sidak_social_confint$lower.CL, digits = 5),
                upper_cl = round(sidak_social_confint$upper.CL, digits = 5),
                outcome = syl_char$outcome[match(sidak_social_confint$name, syl_char$name)])
sidak_social <- summary(pairs(mm_social, adjust = 'sidak')) # tratio and p-val for stats
tests$`t-ratio` <- round(sidak_social$t.ratio, digits = 5)
tests$`p-value` <- round(sidak_social$p.value, digits = 5)

em_plot <- ggplot(tests, aes(name, estimate, color = name)) + 
                  geom_hline(yintercept = 0, color='gray30') + 
                  geom_point(size = 3)  + 
                  geom_errorbar(aes(ymin = lower_cl, ymax = upper_cl), width = 0.4, linewidth = 1.1) + 
                  facet_wrap(outcome ~ ., scales = 'free_x') + 
                  scale_color_manual(values = c('#030637','#030637','#030637',
                                              '#030637','#E95793','gray30')) + # Color based on direction of effect
                  scale_fill_manual(values = c('#030637','#030637','#030637',
                                              '#030637','#E95793','gray30')) + # Color based on direction of effect
                  scale_y_continuous(breaks = pretty_breaks(), limits = c(-1.5,2.5),
                                     name = 'Estimated Marginal Mean Difference') + 
                  plot_theme + 
                  theme(axis.title.x = element_blank(),
                        axis.text.x = element_text(angle = 45, hjust = 1),
                        legend.position = 'none')
```

```{r}
#| label: fig 4a plot
#| echo: false 
#| fig-width: 8
#| fig-asp: 0.75
#| fig-align: center

em_plot
kable(lme_syl_char_table,
      col.names = c('','Value','Std. Error', 'DF', 't-Value'),
      caption = 'LME Social Identity * name')
kable(tests, caption = 'Sidak Post-Hoc Contrasts: Down - Up')
```

### $\Delta$Velocity correlates with PC2 and the SMS (Fig. 4b)

```{r}

```

### $\Delta$Velocity and the SMS detect stress effects equally well (Fig. 4c, 4d)

```{r}

## Distributions Fig. 4c
format <- scale_color_manual(values = col_full[c(1,3,2,4)]) + 
          scale_fill_manual(values = col_full[c(1,3,2,4)]) + 
          plot_theme + 
          theme(legend.position = 'none',
                axis.line.x = element_blank(),
                axis.text.x = element_blank(), 
                axis.title.x = element_blank(),
                axis.ticks.x = element_blank())

SI_distr <- ggplot() + 
              geom_density(data = SIT_diff[SIT_diff$Sex == 'Female', ], 
                           aes(x = IZ_time, color = full_cond, fill = full_cond),
                           alpha = 0.4) + 
              geom_density(data = SIT_diff[SIT_diff$Sex == 'Male', ], 
                           aes(x = IZ_time, y = -(..density..), color = full_cond, fill = full_cond),
                           alpha = 0.4) + 
              coord_flip() + xlab('SI Ratio') +
              ylim(c(-1.5,1.2)) + format

diff_distr <- ggplot() + 
                geom_density(data = SIT_diff[SIT_diff$Sex == 'Female', ], 
                             aes(x = Diff_score, color = full_cond, fill = full_cond),
                             alpha = 0.4) + 
                geom_density(data = SIT_diff[SIT_diff$Sex == 'Male', ], 
                             aes(x = Diff_score, y = -(..density..), color = full_cond, fill = full_cond),
                             alpha = 0.4) + 
                coord_flip() + xlab('SMS') + 
                ylim(c(-1.5,1.5)) + format

vel_distr <- ggplot() + 
              geom_density(data = SIT_diff[SIT_diff$Sex == 'Female', ], 
                           aes(x = Velocity, color = full_cond, fill = full_cond),
                           alpha = 0.4) + 
              geom_density(data = SIT_diff[SIT_diff$Sex == 'Male', ], 
                           aes(x = Velocity, y = -(..density..), color = full_cond, fill = full_cond),
                           alpha = 0.4) + 
              coord_flip() + xlab(expression(Delta*' Velocity'))
              ylim(c(-0.35,0.35)) + format

distr <- plot_grid(SI_distr, diff_distr, vel_distr,
                   ncol = 3, nrow = 1, align = 'hv')

## Effect sizes Fig. 4d

# Calculate abs. ESs by cohort
SIT_diff %>%
  group_by(cohort, Sex) %>%
  summarise(g_SI_ratio = abs(effectsize::hedges_g(IZ_time ~ Condition, .data)$Hedges_g),
            g_diff_score = abs(effectsize::hedges_g(Diff_score ~ Condition, .data)$Hedges_g),
            g_Vel_diff = abs(effectsize::hedges_g(Velocity ~ Condition, .data)$Hedges_g)) %>%
  pivot_longer(cols = c('g_SI_ratio', 'g_diff_score', 'g_Vel_diff'),
               values_to = 'hedges_g', names_to = 'metric') %>%
  mutate(metric = factor(metric, levels = c('g_SI_ratio', 'g_diff_score', 'g_Vel_diff'),
                           labels = c('SI Ratio', 'Diff. Score', 'Velocity')),
         Sex = factor(Sex, levels = c('Male', 'Female'))) -> g_es_cohort

# Calculate overall ES
SIT_diff %>%
  group_by(Sex) %>%
  summarise(abs(g_SI_ratio = effectsize::hedges_g(IZ_time ~ Condition, .data)$Hedges_g),
            abs(g_diff_score = effectsize::hedges_g(Diff_score ~ Condition, .data)$Hedges_g),
            abs(g_Vel_diff = effectsize::hedges_g(Velocity ~ Condition, .data)$Hedges_g)) %>%
  pivot_longer(cols = c('g_SI_ratio', 'g_diff_score', 'g_Vel_diff'),
               values_to = 'hedges_g', names_to = 'metric') %>%
  mutate(metric = factor(metric, levels = c('g_SI_ratio', 'g_diff_score', 'g_Vel_diff'),
                           labels = c('SI Ratio', 'Diff. Score', 'Velocity')),
         Sex = factor(Sex, levels = c('Male', 'Female'))) -> g_es

# Plot
es_plot <- ggplot(g_es_cohort, aes(Sex, hedges_g)) + 
                  geom_hline(yintercept = pwr.t.test(n = 100, sig.level = 0.05, power = 0.8)$d, 
                             color = 'gray80') +
                  geom_hline(yintercept = pwr.t.test(n = 25, sig.level = 0.05, power = 0.8)$d, 
                             color = 'gray80') + 
                  geom_hline(yintercept = pwr.t.test(n = 10, sig.level = 0.05, power = 0.8)$d, 
                             color = 'gray80') + 
                  geom_jitter(height = 0, width = 0.08, size = 2, alpha = 0.5) + 
                  stat_summary(fun = 'mean', geom = 'point', size = 3) + 
                  stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                               linewidth = 1, width = 0.35) + 
                  ylab(expression("Absolute Hedge's"*" "*italic("g"))) + 
                  facet_grid(. ~ metric) + 
                  plot_theme + 
                  theme(axis.title.x = element_blank(),
                  legend.position = 'none')
```

```{r}
#| label: Fig 4cd plot
#| echo: false
#| fig-width: 10
#| fig-asp: 0.66
#| fig-align: center

plot_grid(distr, es_plot)
```

------------------------------------------------------------------------

### $\Delta$Velocity effects are driven by differences in Target trial between control and stressed females (Fig. S3)

------------------------------------------------------------------------

### Session Info

```{r}
#| label: Session Info
#| echo: False
sessionInfo()
```