---
title: 'Figure 5: $\Delta$Velocity and SI Ratio in other Female Stress Models'
format:
    html:
        code-fold: false
        self-contained: true
        toc: true
editor:
    render-on-save: false
---

In Figure 5 we show that $\Delta$Velocity changes are observed in Chronic Social Non-Discriminatory Defeat (CSNDS) as well as Urine Model Defeat. This code generates Figure 5 as well as statistics for the results section **Exploration of** $\Delta$**Velocity across female social stress models.**

### Libraries & Config

```{r}
#| label: libraries
#| message: false
#| code-fold: true

library(here)
library(glue)
library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
library(cowplot)
library(effectsize)
library(emmeans)
library(pander)
library(knitr)
```

```{r}
#| label: config + paths + data
#| code-fold: true

source(here('styles.R'))
data_path <- here('data')
primary_diff <- readRDS(glue('{data_path}/SIT_diff_primary.rds'))
secondary_diff <- readRDS(glue('{data_path}/SIT_diff_secondary.rds'))

# Bind primary and secondary DFs 
primary_diff %<>%
    filter(Sex == 'Female') %>%
    mutate(model = 'CSW/DS',
           lab = 'bagot',
           other = NA) %>%
    select(names(secondary_diff)) 
SIT_diff <- bind_rows(primary_diff, secondary_diff) %>%
                      mutate(other = ifelse(is.na(other), 'None', other))

```

---

### SI ratio across female defeat models (Fig. 5a-d)

SI ratio is significantly lower in stressed females Urine model defeat, but not following CSNDS or in a validation cohort of CSW/DS. In male witnesses, SI ratio is not significantly reduced either. In CSNDS males and in the males of the CSW/DS validation cohort, SI ratio is as expected significantly reduced. 

```{r}
#| label: Fig 5a-d siratio prep
#| code-fold: true

SIT_diff %>%
  filter(model %in% c('CSW/DS VC F', 'CSW/DS VC M','CSW/DS Witness M', 'CSNDS Female', 'CSNDS Male', 'Urine'),
         other != 'ELS') %>%
  mutate(model = factor(model, levels = c('CSW/DS VC F', 'CSW/DS VC M','CSW/DS Witness M', 'CSNDS Female', 'CSNDS Male', 'Urine')), 
         colcond = paste0(Sex, Condition)) -> main_SIT_diff

# PLOT ----
SI_ratio <- ggplot(main_SIT_diff, aes(Condition, IZ_time, color = colcond, fill = colcond)) + 
                geom_hline(yintercept = 1, color = 'gray10') + 
                geom_jitter(width = 0.2, alpha = 0.2, height = 0) + 
                stat_summary(fun = 'mean', geom = 'point', size = 2) + 
                stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                             linewidth = 1, width = 0.4) + 
                facet_grid(. ~ model) + 
                scale_y_continuous(name = 'SI Ratio',
                                   breaks = c(0,1,2,3,4), limits = c(0,4),
                                   expand = c(0,0)) + 
                scale_color_manual(values = col_full) + 
                scale_fill_manual(values = col_full) + 
                plot_theme + 
                theme(legend.position = 'None',
                      axis.title.x = element_blank(),
                      panel.spacing.x = unit(3, 'lines'))

# STATS ----

# ANOVAs
CSWDS_VC_F_data <- SIT_diff[SIT_diff$model == 'CSW/DS VC F',]
aov_SIratio_CSWDS_VC_F <- summary(aov(IZ_time ~ Condition, CSWDS_VC_F_data))

CSWDS_VC_M_data <- SIT_diff[SIT_diff$model == 'CSW/DS VC M',]
aov_SIratio_CSWDS_VC_M <- summary(aov(IZ_time ~ Condition, CSWDS_VC_M_data))

CSWDS_Witness_M_data <- SIT_diff[SIT_diff$model == 'CSW/DS Witness M',] 
aov_SIratio_CSWDS_Witness_M <- summary(aov(IZ_time ~ Condition, CSWDS_Witness_M_data)) 

CSNDS_F_data <- SIT_diff[SIT_diff$model == 'CSNDS Female',]
aov_SIratio_CSNDS_F <- summary(aov(IZ_time ~ Condition, CSNDS_F_data))

CSNDS_M_data <- SIT_diff[SIT_diff$model == 'CSNDS Male',]
aov_SIratio_CSNDS_M <- summary(aov(IZ_time ~ Condition, CSNDS_M_data))

Urine_data <- SIT_diff[SIT_diff$model == 'Urine',]
aov_SIratio_Urine <- summary(aov(IZ_time ~ Condition, Urine_data))

```

```{r}
#| label: Fig 5a-d siratio plot and stats
#| echo: false
#| fig-width: 10
#| fig-asp: 0.6
#| fig-align: center

SI_ratio

pander(aov_SIratio_CSWDS_VC_F, caption = 'ANOVA: SI ratio in in Female CSW/DS witness mice in a validation cohort')
pander(aov_SIratio_CSWDS_VC_M, caption = 'ANOVA: SI ratio in in Male CSW/DS witness mice in a validation cohort')
pander(aov_SIratio_CSWDS_Witness_M, caption = 'ANOVA: SI ratio in Male CSW/DS witness mice')
pander(aov_SIratio_CSNDS_F, caption = 'ANOVA: SI ratio in Female CSNDS mice')
pander(aov_SIratio_CSNDS_M, caption = 'ANOVA: SI ratio in Male CSNDS mice')
pander(aov_SIratio_Urine, caption = 'ANOVA: SI ratio in Urine Defeat')

```


### $\Delta$Velocity across female defeat models (Fig. 5a-d)

As shown in *Figure 4,* $\Delta$Velocity is significantly lower in CSW/DS female witnesses. Here, we demonstrate that $\Delta$Velocity shows a significant reduction in the CSW/DS VC females and following CSNDS, with a trending reduction in Urine model defeat and CSW/DS witness males. The lower effect size after Urine model defeat could relate to the more extreme SI ratio values observed in this model, which are incompatible with low $\Delta$Velocity values, as shown in *Figure 6.*

```{r}
#| label: Fig 5a-d delvel prep
#| warning: false
#| code-fold: true

DelVel <- ggplot(main_SIT_diff, aes(Condition, Velocity, 
                 color = colcond, fill = colcond)) + 
              geom_hline(yintercept = 0, color = 'gray80') + 
              geom_jitter(width = 0.2, alpha = 0.2, height = 0) +
              stat_summary(fun = 'mean', geom = 'point', size = 2) + 
              stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                           linewidth = 1, width = 0.4) + 
              facet_grid(. ~ model) + 
              scale_y_continuous(name = expression(Delta*' Velocity')) + 
              scale_color_manual(values = col_full) + 
              scale_fill_manual(values = col_full) + 
              plot_theme + 
              theme(legend.position = 'None',
                    axis.title.x = element_blank(),
                    panel.spacing.x = unit(3, 'lines'),
                    strip.text = element_blank())

# STATS ----

# ANOVAs
aov_DelVel_CSWDS_VC_F <- summary(aov(Velocity ~ Condition, CSWDS_VC_F_data))
aov_DelVel_CSNDS_VC_M <- summary(aov(Velocity ~ Condition, CSWDS_VC_M_data))
aov_DelVel_CSWDS_Witness_M <- summary(aov(Velocity ~ Condition, CSWDS_Witness_M_data))
aov_DelVel_CSNDS_F <- summary(aov(Velocity ~ Condition, CSNDS_F_data))
aov_DelVel_CSNDS_M <- summary(aov(Velocity ~ Condition, CSNDS_M_data))
aov_DelVel_Urine <- summary(aov(Velocity ~ Condition, Urine_data))

```

```{r}
#| label: Fig 5a-d delvel plot and stats
#| echo: false
#| fig-width: 10
#| fig-asp: 0.6
#| fig-align: center


DelVel

pander(aov_DelVel_CSWDS_VC_F, caption = 'ANOVA: Delta Velocity in in Female CSW/DS witness mice in a validation cohort')
pander(aov_DelVel_CSNDS_VC_M, caption = 'ANOVA: Delta Velocity in in Male CSW/DS witness mice in a validation cohort')
pander(aov_DelVel_CSWDS_Witness_M, caption = 'ANOVA: Delta Velocity in Male CSW/DS witness mice')
pander(aov_DelVel_CSNDS_F, caption = 'ANOVA: Delta Velocity in Female CSNDS mice')
pander(aov_DelVel_CSNDS_M, caption = 'ANOVA: Delta Velocity in Male CSNDS mice')
pander(aov_DelVel_Urine, caption = 'ANOVA: Delta Velocity in Urine Defeat')

```

### Effect sizes across female defeat models (Fig. 5e)

To account for differences in sample size in the datasets used in the study we also provide effect sizes to compare mean differences across different female defeat models. 

```{r}
#| label: Fig 5e stats
#| echo: false
#| warning: false
#| fig-width: 10
#| fig-asp: 0.6
#| fig-align: center

SIT_diff %>%
  filter(model %in% c('CSW/DS VC F', 'CSW/DS VC M','CSW/DS Witness M', 'CSNDS Female', 'CSNDS Male', 'Urine')) %>%
  mutate(model = ifelse(Sex == 'Male' & model == 'CSW/DS', 'CSW/DS Defeat', model)) -> es_SIT_diff

es_SIT_diff %>%
  group_by(model, Sex, other) %>%
  summarise(g_SI_ratio = abs(effectsize::hedges_g(IZ_time ~ Condition, .data)$Hedges_g),
            g_Vel_diff = abs(effectsize::hedges_g(Velocity ~ Condition, .data)$Hedges_g)) %>%
  pivot_longer(cols = c('g_SI_ratio', 'g_Vel_diff'),
               values_to = 'hedges_g', names_to = 'metric') %>%
  mutate(metric = factor(metric, levels = c('g_SI_ratio', 'g_Vel_diff'),
                           labels = c('SI Ratio', 'Velocity'))) %>%
  pivot_wider(values_from = 'hedges_g', names_from = 'metric') -> g_es_models

kable(g_es_models, caption = "Hedge's g effect sizes across models for SI Ratio and Delta Velocity")

```


---

### Models of inter-female aggression (Fig. S5)

#### Mus Muculus (Swiss Webster)
```{r}
#| label: Fig s5a swb
#| echo: false
#| warning: false
#| fig-width: 10
#| fig-asp: 0.6
#| fig-align: center

SIT_diff %>%
  filter(model == 'Swiss Webster') -> swb_SIT_diff

swb_SI <- ggplot(swb_SIT_diff, aes(Condition, IZ_time, color = Condition, fill = Condition)) + 
  geom_hline(yintercept = 1, color = 'gray80') + 
  geom_jitter(aes(shape = other), size = 3, width = 0.2, alpha = 0.2, height = 0) +
  stat_summary(fun.y = 'mean', geom = 'point', size = 3) + 
  stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
               linewidth = 1, width = 0.4) + 
  facet_grid(. ~ model) + 
  ylab('SI Ratio') + 
  scale_color_manual(values = col_full[c(1,2)]) + 
  scale_fill_manual(values = col_full) + 
  plot_theme + 
  theme(legend.position = 'None',
        axis.title.x = element_blank(),
        panel.spacing.x = unit(3, 'lines'))

swb_Vel <- ggplot(swb_SIT_diff, aes(Condition, Velocity, color = Condition, fill = Condition)) + 
  geom_hline(yintercept = 0, color = 'gray80') + 
  geom_jitter(aes(shape = other), size = 3, width = 0.2, alpha = 0.2, height = 0) +
  stat_summary(fun.y = 'mean', geom = 'point', size = 3) + 
  stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
               linewidth = 1, width = 0.4) + 
  facet_grid(. ~ model) + 
  scale_y_continuous(name = expression(Delta*' Velocity'),
                     breaks = c(-2,0,2,4,6,8), limits = c(-2,8)) + 
  scale_color_manual(values = col_full[c(1,2)]) + 
  scale_fill_manual(values = col_full) + 
  plot_theme + 
  theme(legend.position = 'None',
        axis.title.x = element_blank(),
        panel.spacing.x = unit(3, 'lines'))

aov_SIratio_swb <- summary(aov(IZ_time ~ Condition, swb_SIT_diff))
aov_DelVel_swb <- summary(aov(Velocity ~ Condition, swb_SIT_diff))

plot_grid(swb_SI, swb_Vel)

pander(aov_SIratio_swb, caption = 'ANOVA: SI ratio in Mus Muculus (Swiss Webster) model of inter-female aggression.')
pander(aov_DelVel_swb, caption = 'ANOVA: Delta Velocity in Mus Muculus (Swiss Webster) model of inter-female aggression.')
```

#### Peromyscus californicus (California Mouse)

```{r}
#| label: Fig s5b calmouse
#| echo: false
#| fig-width: 10
#| fig-asp: 0.6
#| fig-align: center


SIT_diff %>%
  filter(model == 'California ',
         Sex == 'Female', 
         Velocity > -25 & Velocity < 25) -> cal_SIT_diff

ggplot(cal_SIT_diff, aes(Condition, IZ_time, color = Condition, fill = Condition)) + 
  geom_hline(yintercept = 1, color = 'gray80') + 
  geom_jitter(aes(shape = other), size = 3, width = 0.2, alpha = 0.2, height = 0) +
  stat_summary(fun.y = 'mean', geom = 'point', size = 3) + 
  stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
               linewidth = 1, width = 0.4) + 
  facet_grid(. ~ other) + 
  ylab('SI Ratio') + 
  scale_color_manual(values = col_full[c(1,2)]) + 
  scale_fill_manual(values = col_full) + 
  plot_theme + 
  theme(legend.position = 'None',
        axis.title.x = element_blank(),
        panel.spacing.x = unit(3, 'lines'))

ggplot(cal_SIT_diff, aes(Condition, Velocity, color = Condition, fill = Condition)) + 
  geom_hline(yintercept = 0, color = 'gray80') + 
  geom_jitter(aes(shape = other), size = 3, width = 0.2, alpha = 0.2, height = 0) +
  stat_summary(fun.y = 'mean', geom = 'point', size = 3) + 
  stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
               linewidth = 1, width = 0.4) + 
  facet_grid(. ~ other) + 
  scale_y_continuous(name = expression(Delta*' Velocity')) + 
  scale_color_manual(values = col_full[c(1,2)]) + 
  scale_fill_manual(values = col_full) + 
  plot_theme + 
  theme(legend.position = 'None',
        axis.title.x = element_blank(),
        panel.spacing.x = unit(3, 'lines'))



aov_SIratio_cal <- summary(aov(IZ_time ~ Condition * other, cal_SIT_diff))
aov_SIratio_em <- pairs(emmeans(aov(IZ_time ~ Condition * other, cal_SIT_diff), 
                          ~ Condition | other, adjust = 'sidak'))
aov_DelVel_cal <- summary(aov(Velocity ~ Condition * other, cal_SIT_diff))
aov_DelVel_em <- pairs(emmeans(aov(Velocity ~ Condition * other, cal_SIT_diff), 
                          ~ Condition | other, adjust = 'sidak'))

pander(aov_SIratio_cal, caption = 'ANOVA: SI ratio in P. Californicus model of inter-female aggression.')
kable(aov_SIratio_em, caption = 'Comparisons SI ratio')

pander(aov_DelVel_cal, caption = 'ANOVA: Delta Velocity in P. Californicus model of inter-female aggression.')
kable(aov_DelVel_em, caption = 'Comparisons Delta Velocity')

```

---

## Session Info

```{r}
#| label: session info
#| echo: false

sessionInfo()
```