---
title: 'Figure 5: $\Delta$Velocity and SI Ratio in other Female Stress Models'
format:
    html:
        code-fold: false
        self-contained: true
        toc: true
editor:
    render-on-save: false
---

In Figure 5 we show that $\Delta$Velocity changes are observed in Chronic Social Non-Discriminatory Defeat (CSNDS) as well as Urine Model Defeat. This code generates Fig 5b and Fig 5c as well as statistics for the results section **Decreased** $\Delta$**Velocity generalizes to stressed females in different defeat models**.

### Libraries & Config

```{r}
#| label: libraries
#| message: false
#| code-fold: true

library(here)
library(glue)
library(dplyr)
library(magrittr)
library(ggplot2)
library(effectsize)
library(pander)
library(knitr)
```

```{r}
#| label: config + paths + data
#| code-fold: true

source(here('styles.R'))
data_path <- here('data')
primary_diff <- readRDS(glue('{data_path}/SIT_diff_primary.rds'))
secondary_diff <- readRDS(glue('{data_path}/SIT_diff_secondary.rds'))

# Bind primary and secondary DFs 
primary_diff %<>%
    filter(Sex == 'Female') %>%
    mutate(model = 'CSW/DS',
           lab = 'bagot') %>%
    select(names(secondary_diff)) 
SIT_diff <- bind_rows(primary_diff, secondary_diff) %>%
                mutate(model = factor(model, levels = c('CSW/DS', 'CSNDS', 'Urine')))
```

## Figure 5b

SI ratio is significantly lower in stressed females following CSW/DS and Urine model defeat, but not following CSNDS.

```{r}
#| label: Fig 5b prep
#| code-fold: true

# PLOT ----
SI_ratio <- ggplot(SIT_diff, aes(Condition, IZ_time, color = Condition, fill = Condition)) + 
                geom_hline(yintercept = 1, color = 'gray10') + 
                geom_jitter(width = 0.2, alpha = 0.2, height = 0) + 
                stat_summary(fun = 'mean', geom = 'point', size = 2) + 
                stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                             linewidth = 1, width = 0.4) + 
                facet_grid(. ~ model) + 
                scale_y_continuous(name = 'SI Ratio',
                                   breaks = c(0,1,2,3,4), limits = c(0,4),
                                   expand = c(0,0)) + 
                scale_color_manual(values = col_full[c(1,2)]) + 
                scale_fill_manual(values = col_full) + 
                plot_theme + 
                theme(legend.position = 'None',
                      axis.title.x = element_blank(),
                      panel.spacing.x = unit(3, 'lines'),
                      strip.text = element_blank())

# STATS ----

# ANOVAs & ES
CSWDS_data <- SIT_diff[SIT_diff$model == 'CSW/DS',]
aov_SIratio_CSWDS <- summary(aov(IZ_time ~ Condition + cohort, CSWDS_data))
g_CSWDS <- hedges_g(IZ_time ~ Condition, data = CSWDS_data)$Hedges_g

CSNDS_data <- SIT_diff[SIT_diff$model == 'CSNDS',] 
aov_SIratio_CSNDS <- summary(aov(IZ_time ~ Condition, CSNDS_data)) # Single cohort only
g_CSNDS <- hedges_g(IZ_time ~ Condition, data = CSNDS_data)$Hedges_g

Urine_data <- SIT_diff[SIT_diff$model == 'Urine',]
aov_SIratio_Urine <- summary(aov(IZ_time ~ Condition + cohort, Urine_data))
g_Urine <- hedges_g(IZ_time ~ Condition, data = Urine_data)$Hedges_g

# %Sus
SIT_diff %>%
    group_by(model) %>%
    filter(Condition == 'Stress') %>%
    count(IZ_time < 1) %>%
    mutate(percent_sus = n / sum(n) * 100) %>%
    filter(`IZ_time < 1` == 'TRUE') %>%
    select(!c(`IZ_time < 1`,'n')) %>%
    rename(Model = model, `% Susceptible` = percent_sus) -> pct_sus
pct_sus$`Hedge's g` <- c(g_CSWDS, g_CSNDS, g_Urine)
```

```{r}
#| label: Fig 5b plot and stats
#| echo: false
#| fig-width: 10
#| fig-asp: 0.6
#| fig-align: center

SI_ratio

pander(aov_SIratio_CSWDS, caption = 'ANOVA: SI ratio in CSWDS')
pander(aov_SIratio_CSNDS, caption = 'ANOVA: SI ratio in CSNDS')
pander(aov_SIratio_Urine, caption = 'ANOVA: SI ratio in Urine Defeat')
```

Both CSW/DS and Urine model defeat induce similar effect sizes and produce a similar percentage of susceptible females, with a slightly larger effect observed in Urine model defeat.

```{r}
#| label: Fig 5b extra stats
#| echo: false

kable(pct_sus)
```

## Figure 5c

As shown in *Figure 4,* $\Delta$Velocity is significantly lower in CSW/DS stressed females. Here, we demonstrate that $\Delta$Velocity shows a significant and trending reduction following CSNDS and Urine model defeat, respectively. The lower effect size after Urine model defeat could relate to the more extreme SI ratio values observed in this model, which are incompatible with low $\Delta$Velocity values, as shown in *Figure 6.*

```{r}
#| label: Fig 5c prep
#| warning: false
#| code-fold: true

DelVel <- ggplot(SIT_diff, aes(Condition, Velocity, color = Condition, fill = Condition)) + 
              geom_hline(yintercept = 0, color = 'gray80') + 
              geom_jitter(width = 0.2, alpha = 0.2, height = 0) +
              stat_summary(fun = 'mean', geom = 'point', size = 2) + 
              stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                           linewidth = 1, width = 0.4) + 
              facet_grid(. ~ model) + 
              scale_y_continuous(name = expression(Delta*' Velocity')) + 
              scale_color_manual(values = col_full[c(1,2)]) + 
              scale_fill_manual(values = col_full) + 
              plot_theme + 
              theme(legend.position = 'None',
                    axis.title.x = element_blank(),
                    panel.spacing.x = unit(3, 'lines'),
                    strip.text = element_blank())

# STATS ----

# ANOVAs
aov_SIratio_CSWDS <- summary(aov(Velocity ~ Condition + cohort, CSWDS_data))
aov_SIratio_CSNDS <- summary(aov(Velocity ~ Condition, CSNDS_data)) # Single cohort only
aov_SIratio_Urine <- summary(aov(Velocity ~ Condition + cohort, Urine_data))

```

```{r}
#| label: Fig 5c plot and stats
#| echo: false
#| fig-width: 10
#| fig-asp: 0.6
#| fig-align: center


DelVel

pander(aov_SIratio_CSWDS, caption = 'ANOVA: Delta Velocity in CSWDS')
pander(aov_SIratio_CSNDS, caption = 'ANOVA: Delta Velocity in CSNDS')
pander(aov_SIratio_Urine, caption = 'ANOVA: Delta Velocity in Urine Defeat')
```

## Session Info

```{r}
#| label: session info
#| echo: false

sessionInfo()
```