---
title: 'Figure 5: $\Delta$Velocity and SI Ratio in other Female Stress Models'
format:
    html:
        code-fold: false
        self-contained: true
        toc: true
editor:
    render-on-save: false
---

In Figure 5 we show that $\Delta$Velocity changes are observed in Chronic Social Non-Discriminatory Defeat (CSNDS) as well as Urine Model Defeat. This code generates Fig 5b and Fig 5c as well as statistics for the results section **Decreased** $\Delta$**Velocity generalizes to stressed females in different defeat models**.

### Libraries & Config

```{r}
#| label: libraries
#| message: false
#| code-fold: true

library(here)
library(glue)
library(dplyr)
library(ggplot2)

```

```{r}
#| label: config + paths + data
#| code-fold: true

source(here('styles.R'))
data_path <- here('data')
primary_diff <- readRDS(glue('{data_path}/SIT_diff_primary.rds'))
secondary_diff <- readRDS(glue('{data_path}/SIT_diff_secondary.rds'))

# Bind primary and secondary DFs 
SIT_diff <- bind_rows(primary_diff, secondary_diff)
```

## Figure 5b

SI ratio is significantly lower in stressed females following CSW/DS and Urine model defeat, but not following CSNDS. Both CSW/DS and Urine model defeat induce similar effect sizes and produce a similar percentage of susceptible females.

```{r}
#| label: Fig 5b prep
#| code-fold: true
```

## Figure 5c

## Session Info

```{r}
#| label: session info
#| echo: false

sessionInfo()
```