---
title: 'Figure 7: mPFC->NAc FIP'
format:
    html:
        code-fold: false
        self-contained: true
        toc: true
editor:
    render-on-save: false
---

Figure 7 demonstrates that a classification based on $\Delta$Velocity *and* SI ratio identifies meaningful neurobiological differences between resilient and susceptible female mice. This notebook generates Fig 7 as well as statistics for the results section **Social modulation of mPFC-NAc neural activity in resilient, but not susceptible, females**.

### Libraries & Config

```{r}
#| label: libraries
#| message: false
#| code-fold: true

library(here)
library(glue)
library(emmeans)
library(nlme)
library(ggplot2)
library(knitr)
library(pander)
```

```{r}
#| label: config + paths + data
#| code-fold: true

source(here('styles.R'))
data_path <- here('data')
FIP_data <- readRDS(glue('{data_path}/FIP_data.rds'))
```

---

### Neural activity is modulated by target presence arena-wide in resilient females only (Fig. 7b)

We show that FIP recordings in the mPFC to NAc pathway in resilient mice is different in the No Target and Target trial across the whole arena. In contrast, susceptible mice do not show differences between No Target and Targe trials, which matches the observation of non-adaptation on the behavioral level. Controls also do not show discrimination between trial types on the level of neuronal recordings.

```{r}
#| label: Fig 7b prep
#| code-fold: true

# PLOT ----
FIP_Arena <- ggplot(FIP_data, aes(Trial_type, z_Arena)) + 
                      geom_hline(yintercept = 0) + 
                      geom_jitter(
                        data = FIP_data[FIP_data$Group_combined != 'ResSus',],
                        aes(alpha = Trial_type, color = Group), 
                        width = 0.2, size = 5.5, height = 0) + 
                      geom_jitter(
                        data = FIP_data[FIP_data$Group_combined == 'ResSus',],
                        aes(alpha = Trial_type), color = '#C1292E', 
                        shape = 17, width = 0.2, size = 5.5, height = 0) + 
                      stat_summary(fun = 'mean', geom = 'point', size = 5, 
                                   color = 'gray50') + 
                      stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                                   linewidth = 1.5, width = 0.5, color = 'gray50') + 
                      facet_grid(. ~ Group) + 
                      ylim(c(-1,.5)) + ylab('Mean Z-Score') +
                      plot_theme + 
                      scale_color_manual(values = col_stress_f) + 
                      scale_alpha_manual(values = c(0.8,0.5)) + 
                      theme(axis.title.x = element_blank(),
                            legend.position = 'none')

# STATS ----

lme_arena_f <- lme(z_Arena ~ Group * Trial_type, 
                     data = FIP_data, random = list(applied_ID = ~ 1))
lme_arena_f_table <- cbind(rownames(summary(lme_arena_f)$tTable),
                            unname(summary(lme_arena_f)$tTable[,'Value']),
                            unname(summary(lme_arena_f)$tTable[,'Std.Error']),
                            unname(summary(lme_arena_f)$tTable[,'DF']),
                            unname(summary(lme_arena_f)$tTable[,'t-value']))
mm_ttype_arena_f <- pairs(emmeans(lme_arena_f, ~ Trial_type | Group), adjust = 'sidak')
```

```{r}
#| label: Fig 7b plot and stats
#| echo: false
#| fig-width: 7
#| fig-asp: 0.75
#| fig-align: center

FIP_Arena

kable(lme_arena_f_table, 
      col.names = c('','Value','Std. Error', 'DF', 't-Value'),
      caption = 'LME in Whole Arena')
kable(mm_ttype_arena_f, align='c', caption = 'Sidak Post-Hoc Planned Contrasts in Whole Arena')
```

### Neural activity is modulated by target presence in the IZ in resilient females only (Fig. 7c)

We further show that the same effect when animals are in the interaction zone.

```{r}
#| label: Fig 7c prep
#| code-fold: true

# PLOT ----
FIP_IZ <- ggplot(FIP_data, aes(Trial_type, z_IZ)) + 
                      geom_hline(yintercept = 0) + 
                      geom_jitter(
                        data = FIP_data[FIP_data$Group_combined != 'ResSus',],
                        aes(alpha = Trial_type, color = Group), 
                        width = 0.2, size = 5.5, height = 0) + 
                      geom_jitter(
                        data = FIP_data[FIP_data$Group_combined == 'ResSus',],
                        aes(alpha = Trial_type), color = '#C1292E', 
                        shape = 17, width = 0.2, size = 5.5, height = 0) + 
                      stat_summary(fun = 'mean', geom = 'point', size = 5, 
                                   color = 'gray50') + 
                      stat_summary(fun.data = 'mean_se', geom = 'errorbar', 
                                   linewidth = 1.5, width = 0.5, color = 'gray50') + 
                      facet_grid(. ~ Group) + 
                      ylim(c(-1.5,.5)) + ylab('Mean Z-Score') +
                      plot_theme + 
                      scale_color_manual(values = col_stress_f) + 
                      scale_alpha_manual(values = c(0.8,0.5)) + 
                      theme(axis.title.x = element_blank(),
                            legend.position = 'none')

# STATS ----

lme_IZ_f <- lme(z_IZ ~ Group * Trial_type, 
                     data = FIP_data, random = list(applied_ID = ~ 1))
lme_IZ_f_table <- cbind(rownames(summary(lme_IZ_f)$tTable),
                          unname(summary(lme_IZ_f)$tTable[,'Value']),
                          unname(summary(lme_IZ_f)$tTable[,'Std.Error']),
                          unname(summary(lme_IZ_f)$tTable[,'DF']),
                          unname(summary(lme_IZ_f)$tTable[,'t-value']))
mm_ttype_IZ_f <- pairs(emmeans(lme_IZ_f, ~ Trial_type | Group), adjust = 'sidak')
```

```{r}
#| label: Fig 7c plot and stats
#| echo: false
#| fig-width: 7
#| fig-asp: 0.75
#| fig-align: center

FIP_IZ

kable(lme_IZ_f_table, 
      col.names = c('','Value','Std. Error', 'DF', 't-Value'),
      caption = 'LME in Interaction Zone', )
kable(mm_ttype_IZ_f, align='c', caption = 'Sidak Post-Hoc Planned Contrasts in Interaction Zone')
```

---

### Session Info

```{r}
#| label: Session Info
#| echo: False
sessionInfo()
```